From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-29 Pi+/paper/finite.tex
More progress on sec 6

We already mentioned that the Coxter presentation, which is the one will use, is based on adjacent transpositions, meaning that the primitive operations we use are adjacent swaps. However, $\Sn$ is not a free group, and so we also have to specify the relations of the presentaiton -- the laws that such generators have to satisfy. there are three of them -- and it is easiest to visualise them as braid diagrams.

Swapping the same two elements two times in a row is the same as doing nothing:

@@

When swapping two distinct pairs of elements (i.e. when the indices of two transpositions
differ by at least 1), it does not matter in which order the swapping happens,
that is, we can slide the wires freely.

@@

This construction is called a Coxeter presentation. Writing it formally, we define a binary relation $\cox$ on
$\List[\Fin[n]]$.

@@

@@

Generators $\longcancel$ and $\longswap$ correspond directly to the generators of $\cox*$ and can be visualised in the
same way as before. The generator $\longbraid$ uses a helper $n \downf k$, which produces the sequence
\(k + n, k + n - 1 \ldots, n + 1\).

>>>>

We already mentioned that the Coxter presentation, which is the one will use, is based on adjacent transpositions, meaning that the primitive operations we use are adjacent swaps. When dealing with permutations on an $n + 1$-element set, there are $n$ adjacent transpositions, transposition number $k$ swapping elements $\el{k}$ and $\el{k+1}$. Thus, the generating set would be $\Fin[n]$.

However, $\Sn$ is not a free group, and so we also have to specify the relations of the presentaiton -- the laws that such generators have to satisfy. there are three of them -- it is easiest to visualise them as braid diagrams.

First, swapping the same two elements two times in a row is the same as doing nothing:

@@

Second, when swapping two distinct pairs of elements (i.e. when the indices of two transpositions
differ by at least 1), it does not matter in which order the swapping happens,
that is, we can slide the wires freely.

@@

This construction is called a Coxeter presentation. Writing it formally, we take the generating set to be $\Fin[n]$, where the element $k$ corresponds to an adjacent transpositions $\tau_k$, which swaps elements $\el{k}$ and $\el{k+1}$. Then, we define a binary relation $\cox$ on the $\List[\Fin[n]]$, which encodes the laws discussed above.

@@

First, we need a helper function $n \downf k$.

The reuslt of this function is the sequence \(k + n, k + n - 1, k + n - 2, \ldots, n + 1\). If we think of the element $m$ as the transposition which swaps elements $\el{m}$ and $\el{m+1}$, the role of this helper function is to produce a sequence of transpositions - a permutation - which moves element $n$ $k$ places right.

@@

Generators $\longcancel$ and $\longswap$ correspond directly to the generators of $\cox*$ and can be visualised in the
same way as before. The generator $\longbraid$
